---
title: "Live session Unit 05 assignment"
author: "Matt Pavlovich"
date: "10/01/2017"
output: 
  html_document:
      keep_md: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# set the path to your data files here
knitr::opts_knit$set(root.dir = "~/")
```
# Link to homework files:
#### https://github.com/meacreatio/SMU_MSDS_Homework/tree/master/Homework/6306/week5

# Question 1: Data Munging 
#### a. Import text file and rename colmns
```{r}
df <- read.table("yob2016.txt", sep = ";", header = F)
names(df) <- c("Name", "Sex", "Count")
```

#### b. Data Summary and Structure
```{r}
summary(df)
str(df)
dim(df)
```

#### c. Display name that ends with 'yyy'
```{r}
df.mispelled.name <- df[grep("yyy$", df$Name), ]
as.character(df.mispelled.name$Name[1])
```

#### d. Remove invalid name and create new dataframe reference
```{r}
index.misspelled.name <- which(df$Name == df.mispelled.name$Name[1])
y2016 <- df[-index.misspelled.name, ]
```

# Question 2: Data Merging
#### a. Import text file and rename colmns
```{r}
y2015 <- read.table("yob2015.txt", sep = ",", header = F)
names(y2015) <- c("Name", "Sex", "Count")
```

#### b. Display last ten rows
```{r}
tail(y2015, n = 10)
# all names are male and the dataset is ordered by sex ascending and count descending
```

#### c. Merge 2016 and 2015
```{r}
final <- merge(y2015, y2016, by = c("Name", "Sex"), all = T)
final <- final[complete.cases(final), ]
```

# Question 3: Data Summary
#### a. Amount of children
```{r}
final$Total <- rowSums(final[, c("Count.x", "Count.y")])
sum(final$Total)
```

#### b. Sort by Total and display top ten names
```{r}
df.names.sorted.total <- final[order(-final$Total), ]
head(df.names.sorted.total, n = 10)
```

##### c. Display top ten girls' names
```{r}
df.names.sorted.total.girls <- df.names.sorted.total[df.names.sorted.total$Sex == "F", ]
girl <- df.names.sorted.total.girls[1:10, ]
girl
```

#### d. Write top 10 girls names to a csv file
```{r}
write.csv(girl[, c("Name", "Total")], file = "itsagirl.csv", row.names = F)
```
